---
title: "Problem 7 - Regression: California Housing"
author: "Henry Neeb, Christopher Kurrus, Tyler Chase, and Yash Vyas"
date: "May 22, 2016"
output: pdf_document
---

## Libraries

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(gbm)
```

## File Parameters

```{r}
# Session -> set working directory -> Source file location

set.seed(131)

# Path for data
data_path <- "data/"
data_name <- "California_Data.txt"
data_name <- paste(data_path, data_name, sep = "")

```

## Read in Data

```{r}
data_0 <- read_csv(file = data_name, col_names = FALSE)


# Name data columns
column_names <- c("House_Value", "Median_Income", "Housing_Median_Age",
                  "Average_Rooms", "Average_Bedrooms", "Population",
                  "Average_Occupancy", "Lattitude", "Longitude")
colnames(data_0) <- column_names

# Make sure the samples are mixed
set.seed(131)
data <- data_0[sample(nrow(data_0)),]
```

## Part (a)

We fit our data using a gbm model and determine the prediction accuracy. 

```{r}
# Fit model on our data
set.seed(444) # Random for bag.fraction
gbm_model <- gbm(House_Value ~ ., data = data , train.fraction = 0.8, interaction.depth = 4,
                shrinkage = 0.05, n.trees = 10000, bag.fraction = 0.5, cv.folds = 5,
                distribution = "gaussian", verbose = F)

# Determine the optimal number of trees by cross validation 
best.iter <- gbm.perf(gbm_model, method = "cv")
```

Determine the training error and the test error. 

```{r}
# Determine the training error and the test error
train_error = gbm_model$train.error[best.iter]
test_error = gbm_model$valid.error[best.iter]

```
We quantify the accuracy of the gbm on the data by reporting the mean square error for both the training set and the test set. The training error is `r train_error`, while the test error is `r test_error`. 

## Part (b)

Here we identify the most important variables. 

```{r}
summary(gbm_model, main = "Relative Influence of all Predictors")
```

I will list the 4 variables that are over 10% in relative influence. Median income is the most influential variable with a relative influence of 44.1. Average Occupancy is the most second most influencial is the average occupancy which has a relative influence of 14.3. The third most influencial is the longitude with a relative influence of 12.9, followed by lattitude with a relative influence of 10.7. 

## Part (c)

Here we plot the dependence of the response on the most important variables, and on the most important variable pairs. 


```{r}
plot(x=gbm_model, i.var=1, n.trees=best.iter)
plot(x=gbm_model, i.var=6, n.trees=best.iter)
plot(x=gbm_model, i.var=7, n.trees=best.iter)
plot(x=gbm_model, i.var=8, n.trees=best.iter)
plot(x=gbm_model, i.var=c(1,6), n.trees=best.iter)
plot(x=gbm_model, i.var=c(1,7), n.trees=best.iter)
plot(x=gbm_model, i.var=c(1,8), n.trees=best.iter)
plot(x=gbm_model, i.var=c(6,7), n.trees=best.iter)
plot(x=gbm_model, i.var=c(7,8), n.trees=best.iter)
plot(x=gbm_model, i.var=c(6,8), n.trees=best.iter)
```

We can see that median income has a positive correlation with house value. Average occupancey has a negative correlation with house value and has a steep drop at low occupancy. Lattitude has a netative correlation with house value. And Longitude also has a negative correlation with house value. 

Occupancy and median income don't seem very correlated. There does seem to be a correlation between Lattitude and median income. The lower the lattitude and higher the median income the more the house value. There is also a correlation between longitude and median income. The lower the longitude and higher the median income the more the house value. There doesn't seem to be a correlation between Latitude and average occupancy. There does seem to be a correlation between longitude and lattitude. The lower the lattitude and longitude the more the house value. 