---
title: "Probelem 6 - Binary Classification: Spam Email"
author: "Henry Neeb, Christopher Kurrus, Tyler Chase, and Yash Vyas"
date: "May 22, 2016"
output: pdf_document
---

## Libraries

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(gbm)
```

## File Parameters

```{r}
# Session -> set working directory -> Source file location

# Path for data
data_path <- "data/"
file <- "Income_Data.txt"
ext <- paste(data_path, file, sep = "")
```

## Read in Data

The data does not have any column headers, so we have to manually assign the 
column names. Our variables are also all categorical variables, some with an 
order relationship and some with out. We also have to import these values as 
factors, specifying if they have an order relationship or not. The following 
variables have an order relationship:

* Income
* Age
* Education
* Amount of time living in the bay area
* Household count
* Minors in household count

The remaining variables do not have an order relationship:

* Sex
* Marital status
* Occupation
* Dual income status
* Whether you rent or own a house
* Type of house you live in
* Ethnicity
* Languages

```{r, cache=TRUE}
# Name of the variables imported
var_names = c("income",
              "sex",
              "marital_status",
              "age",
              "education",
              "occupation",
              "bay_duration",
              "dual_income",
              "household_total",
              "household_minors",
              "householder_status",
              "home_type",
              "ethnicity",
              "language")

# Type of the variable being imported
var_type <- cols(col_factor(sprintf("%i", 1:9), ordered =  TRUE), # income
                 col_factor(sprintf("%i", 1:2), ordered = FALSE), # sex
                 col_factor(sprintf("%i", 1:5), ordered = FALSE), # marital stat
                 col_factor(sprintf("%i", 1:7), ordered =  TRUE), # age
                 col_factor(sprintf("%i", 1:6), ordered =  TRUE), # education
                 col_factor(sprintf("%i", 1:9), ordered = FALSE), # occupation
                 col_factor(sprintf("%i", 1:5), ordered =  TRUE), # bay duration
                 col_factor(sprintf("%i", 1:3), ordered = FALSE), # dual income
                 col_factor(sprintf("%i", 1:9), ordered =  TRUE), # house count
                 col_factor(sprintf("%i", 0:9), ordered =  TRUE), # minor count
                 col_factor(sprintf("%i", 1:3), ordered = FALSE), # rent/own
                 col_factor(sprintf("%i", 1:5), ordered = FALSE), # house type
                 col_factor(sprintf("%i", 1:8), ordered = FALSE), # ethnicity
                 col_factor(sprintf("%i", 1:3), ordered = FALSE)) # language

df <- read_csv(file = ext, 
               col_names = var_names, 
               col_types = var_type)
```

## Randomize the Dat

To use GBM's test/train functionality, we will randomize the data first. We do 
this because GBM will take the first x% specified as training data and the 
remaining as testing data. We want to ensure that the data is properly shuffled 
first to mimic random sampling across our data.

```{r}
# Randomize the data
n <- dim(df)[1]
df <- df[sample(n),]

# Split 80% training and 20% test
train_n <- round(0.8 * n, 0)
test_n <- n - train_n
train <- df[train_n,]
test <- df[test_n,]
```

## Fitting the GBM

We now fit our first model. We will make our first model the same as that in 
the tutorial.

```{r}
gbm0 <- gbm(income~., 
            data = df, 
            train.fraction = 1.0, 
            interaction.depth = 1, 
            shrinkage = .05, 
            n.trees = 2500, 
            bag.fraction = 0.5, 
            cv.folds = 5, 
            distribution = "multinomial", 
            verbose = TRUE)

gbm.perf(object = gbm0, method = "cv")
bestIter_test <- gbm.perf(object = gbm0, method = "cv")
summary(gbm0)
```

We use 5 fold cross validation to pick our best number of trees. We will then 
use this amount (`bestIter_test`) to predict on our holdout testing data.

```{r}

```


With a direct copy from the tutorial and using a training/testing method for 
selecting the correct number of iterations, we pick to use `bestIter_test` as 
our optimal number of iterations.


Things to test:

* Shrinkage
* bag.fraction
* interaction.depth
* shrinkage
