---
title: "Problem 6 - Binary Classification: Spam Email"
author: "Henry Neeb, Christopher Kurrus, Tyler Chase, and Yash Vyas"
date: "May 22, 2016"
output: pdf_document
---

## Libraries

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(gbm)
```

## File Parameters

```{r}
# Session -> set working directory -> Source file location

set.seed(131)

# Path for data
data_path <- "data/"
file_tst <- "Spam.Test.txt"
file_trn <- "Spam_Train.txt"
file_tst <- paste(data_path, file_tst, sep = "")
file_trn <- paste(data_path, file_trn, sep = "")
```

## Read in Data

```{r}
tst <- read_csv(file = file_tst, col_names = FALSE)
trn <- read_csv(file = file_trn, col_names = FALSE)
```

Here we require both the testing and training spam datasets from the course website, stored as comma separated values.

To aid in interpretability later we will correct the column names now.

```{r}
cnms <- c("make", "address", "all", "3d", "our", "over", "remove", "internet","order", "mail", "receive", "will", "people", "report", "addresses","free", "business", "email", "you", "credit", "your", "font","000","money", "hp", "hpl", "george", "650", "lab", "labs", "telnet", "857", "data", "415", "85", "technology", "1999", "parts","pm", "direct", "cs", "meeting", "original", "project", "re","edu", "table", "conference", ";", "(", "[", "!", "$", "#", "CAPAVE", "CAPMAX", "CAPTOT","type") 

colnames(tst) <- cnms
colnames(trn) <- cnms

```


## Part (a)

Here we will fit a gbm model on the training data and check its misclassification on the testing dataset.

```{r}

gbm0 <- gbm(type ~ . , data = trn, train.fraction = 1, interaction.depth=4, shrinkage=.05,  n.trees=2500, bag.fraction=0.5, cv.folds=5, distribution="bernoulli", verbose=T)

gbm0.pred <- predict(gbm0, tst, type="response", n.trees=300)
df <- cbind(as.data.frame(gbm0.pred), tst$type)
df <- df %>% mutate(pred = ifelse(gbm0.pred >= .5, 1, 0), equ = pred == tst$type)
gbm0.tst.mc <- (1534 - sum(df$equ))/1534
gbm0.tst.mc

df_spam <- df %>% filter(tst$type == 1)
gbm0.spam.mc <- (618 - sum(df_spam$equ))/618
gbm0.spam.mc

df_not <- df %>% filter(tst$type == 0)
gbm0.not.mc <- (916 - sum(df_not$equ))/916
gbm0.not.mc
```

Here we can see our overall estimate of misclassification rate is .0391.  Our misclassification percentage for spam in the testing set is 5.18% and for not spam is 3.06%


## Part (B)

Now we will set up a function to let us quickly edit our cost parameter and find our misclassification rates.

```{r}

cost <- function(cst) {
df <- cbind(as.data.frame(gbm0.pred), tst$type)
df <- df %>% mutate(pred = ifelse(gbm0.pred >= cst, 1, 0), equ = pred == tst$type)
gbm0.tst.mc <- (1534 - sum(df$equ))/1534
gbm0.tst.mc

df_spam <- df %>% filter(tst$type == 1)
gbm0.spam.mc <- (618 - sum(df_spam$equ))/618
gbm0.spam.mc

df_not <- df %>% filter(tst$type == 0)
gbm0.not.mc <- (916 - sum(df_not$equ))/916
gbm0.not.mc

return(c(gbm0.tst.mc, gbm0.spam.mc, gbm0.not.mc))
}

cost(.5)
cost(.6)
cost(.8)
cost(.955)

```

Now we see that with a very high relative cost we can reach the desired non-spam misclassification rate, but our overall misclassification rate has risen to .1545.  Our misclassification percentage for spam in the testing set is now 38.02% and for not spam is .22%

```{r}

summary(gbm0,main="RELATIVE INFLUENCE OF ALL PREDICTORS")
best.iter<-gbm.perf(gbm0,method="OOB")#Best iteration by OOB 
#plot(x=gbm0,i.var=58,n.trees=best.iter, main="Partial Dependence of 'type' ") 


```



 








